### 2019-09-25 ch9_discussion/content.tex
update

\section{Self-assessment in trajectory inference}
Many trajectory inference (TI) tools lack quantitative assessment of the accuracy of the method. Instead, they rely on anecdotal evidence to demonstrate their added value. A brief review of 76 TI articles reveals that only about 35\% contain a self-assessment (Figure \ref{fig:benchmarks_over_time}). Peer-reviewed TI articles fared even worse, self-assessing in only 32\% of cases (n=52), whereas articles first published as a pre-print self-assess in 42\% of cases (n=40). Only four TI articles feature a decently-sized comparison of at least 5 methods using 5 datasets or more.
% Note: i did not count how many metrics each of these manuscripts used.

@@

Several authors benchmark their method against synthetic data in order to evaluate their method \cite{haghverdi_diffusionmapshighdimensional_2015, street_slingshotcelllineage_2018,rizvi_singlecelltopologicalrnaseq_2017,welch_matchermanifoldalignment_2017,guo_hoplandsinglecellpseudotime_2017,boukouvalas_bgpbranchedgaussian_2017,gao_calistaclusteringlineage_2018,parra_reconstructingcomplexlineage_2018,chen_singlecelltrajectoriesreconstruction_2019,strauss_gpseudoclustdeconvolutionshared_2019}. These simulators are mostly custom made by the authors themselves, thereby increasing the level of bias in the benchmark. Instead, readers are encouraged to use of the readily available simulators, namely dyntoy \cite{saelens_comparisonsinglecelltrajectory_2019}, dyngen \cite{dyngen}, splatter \cite{zappia_splattersimulationsinglecell_2017}, and PROSSTT \cite{papadopoulos_prossttprobabilisticsimulation_2018}.

@@

Several authors use mainly real datasets to evaluate their method, though only few use more than four datasets\cite{cannoodt_scorpiusimprovestrajectory_2016,liu_reconstructingcellcycle_2017,sharma_forksfindingorderings_2017,guo_hoplandsinglecellpseudotime_2017,teschendorff_singlecellentropyaccurate_2017,macnair_psupertimesupervisedpseudotime_2019}. 
By now, already hundreds of suitable real datasets exist (Figure~\ref{fig:datasets}). Downloading and pre-processing them requires a significant time investment, but by processing the datasets once and storing them in a single repository they can be reused for multiple purposes.

>>>>

\section{Self-assessment in trajectory inference}
Many trajectory inference (TI) tools lack quantitative assessment of the accuracy of the method. Instead, they rely on anecdotal evidence to demonstrate their added value. A brief review of 75 TI articles reveals that only about 37\% contain a self-assessment (Figure \ref{fig:benchmarks_over_time}). Peer-reviewed TI articles fared even worse, self-assessing in only 34\% of cases (n=55), whereas articles first published as a pre-print self-assess in 43\% of cases (n=39). Only three TI articles feature a comparison of at least 5 methods using 5 datasets or more \cite{sharma_forksfindingorderings_2017,guo_hoplandsinglecellpseudotime_2017,parra_reconstructingcomplexlineage_2018}.

@@

When real data is scarce, synthetic data is often used to evaluate computational methods, either standalone (n=5) or to complement real data (n=7). Most synthetic data is generated by the authors themselves (n=8), whereas some reuse datasets generated by others (n=3) or use one of the readily available simulators (n=2). To avoid introducing self-assessment bias in a benchmark, it is recommended to use readily available simulators if they fit the requirements. Examples are dyntoy \cite{saelens_comparisonsinglecelltrajectory_2019}, dyngen \cite{dyngen}, splatter \cite{zappia_splattersimulationsinglecell_2017}, and PROSSTT \cite{papadopoulos_prossttprobabilisticsimulation_2018}.

@@

Several authors use mainly real datasets to evaluate their method, though only few use more than four datasets (n=7). 
By now, already hundreds of suitable real datasets exist (Figure~\ref{fig:datasets}). Downloading and pre-processing them requires a significant time investment, but by processing the datasets once and storing them in a single repository they can be reused for multiple purposes.
